How to publish in Biological Sciences: a guide for the uninitiated

### 2021-12-15 1.06-authors.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

Who should be an author is being increasingly regulated because of widespread abuse, including ‘honorary authors’ or excluding ‘ghost authors’. Back in the day there were people that used to add fictitious characters to their author line (including their dogs!) as a joke. These days we have such a thing as [ORCID](#orcid) that attempts to register all authors, turning them into numbers ([see part 3](#orcid)). One of the benefits with the ORCID system is that you can have your name the way you want to have it, even if it is not a standard western style surname [@goyesvallejos2021what]. The need for all this regulation is because publications have turned into a kind of currency for academics. And when there is currency involved, abuse quickly results in human systems followed by the need for regulation. Hence, one of the results of going transparent is that we all become registered numbers (and our dogs won't get authorship any more). Perhaps this is just the loss of an age of innocence, at least on the part of our dogs!

@@

## Ghost authorship {#ghost}
Ghost authorship is when people, not included in the author line, have contributed substantially to the study [@matheson2016ghostwriting]. Note that ghostwriting outside science is usually when someone who has the talent to write, writes the ideas of someone who doesn’t for a fee, and in return the latter takes the author line. These are called "paper mills" and they are discussed in [Part I](#papermills)).

@@

### Paper mills {#papermills}
The concept of a paper mill is rather different from [ghost authors](#ghost) or even salami-slicing ([Part IV](#salami)). Paper mills involve third parties, often not included on the author line, producing material for publication from scratch: i.e. companies that specialise in producing content that will pass peer review for those who want to buy authorship. The world of paper mills is particularly shady and it is not clear what sums of money are exchanged for these types of goods. What we do know is that there are benefits to those who are put on the authorline, while those that produce the manuscripts generally rewrite text and pull protocols from manuscripts that are already published. Results are often images that have already been published and/or are manipulated to suit the content (see [Part IV](#fabrication)). Hence, paper mills are a systematic and deliberate manipulation of the publication process [see @teixeiradasilva2021abuse].

## Honorary authorship {#honorary}
Honorary authorship happens when people who have not contributed meaningfully to a study are included in the author line. If publications are the currency of science then you can see how being added to other people's publications increases your apparent productivity. While this might sound surprising to you, you should know that does happen and might be more common than you think. The Wisler et al. [-@wislar2011honorary] study found that honorary authorship was as high as 17.6% in medical publications. It’s worth noting that this may vary between disciplines as there are various traditions in some disciplines whereby the head of a large team may always be included as an author of a paper that emerges from the team, whether or not they were involved. In biological sciences, teams tend to be quite small with a single or rarely multiple Principal Investigators (PIs). This means that your PI will likely be directly involved with your research and therefore also an author. Imagine a very large team with multiple PIs working under a head who insists that they are an author on every publication. This could add up to hundreds of publications in a year (e.g. [Yuri T. Struchkov is currently credited with >1600 publications on Scopus](https://www-scopus-com.ez.sun.ac.za/authid/detail.uri?authorId=35427689100)), and such prolific authorship has been questioned [e.g. @rennie1994authorship]. However, there appear to be very different levels of what could be considered credible and what incredible [e.g. 25 papers a year: @wager2015too], and I respond to this, and the general question of how prolific authors are becoming, [later](#publishmore).

In the biological sciences, the area of molecular phylogenetics has traditionally honoured those who collect tissue samples with authorship on the resulting phylogenies when published. As I mentioned above, this is not always equal and has also been used to politically [honour](#honorary) or [ghost](#ghost). The reason given for honouring contributors in this way is that the studies could not have been done without the tissues. On the other side, some people have long lists of publications based on tissue donations and very little else.

## A need for transparency - DORA {#DORA}
Given the problems with both ghost and honorary authorship, there is clearly a need for transparency about who the authors are and what they have actually contributed. This was recognised by the [Declaration of Research Assessment](DORA: https://sfdora.org/), which has a growing number of signatories as well as some solid ideas on the way forward in assessment of research and researchers. In particular, DORA is against the use of Impact Factors ([see below](#impactfactors)) and other journal based metrics, and instead assesses the research on its own merits. DORA also encourages everyone to embrace the opportunities offered by online publication, including colour figures and unencumbered word lengths. DORA also encourages specific information to be published about individual author contributions. In short, DORA stands for transparency and it would be worth you looking at their statement and finding out whether your institution is a signatory.

@@

## Who should be an author?
There is an increasing number of journals that now give clear instructions on who qualifies as author on a paper, and these have been formalised by the International Committee of Medical Journal Editors (ICMJE), and this area has been rapidly evolving [@baskin2011honorary]. Although these approaches are clearly made to prevent [honorary authorships](#honorary), it is hard to see how they help include [ghosted authors](#ghost). Scientists are notorious in their abilities to make _post hoc_ rationalisations. 
As an early career researcher, I'd suggest that you use questions around authorship as a way to formulate how participating researchers fit into the study. By openly attributing roles to a checklist, everyone buys into what is expected of them. Any additions to this author line-up should be openly discussed amongst all authors, and attibuted to the checklist, as the project proceeds.

>>>>

Who should be an author is being increasingly regulated because of widespread abuse, including ‘honorary authors’ or excluding ‘ghost authors’. Back in the day, there were people that used to add their pets to the author line including their dog, cat, hamster or parrot [@penders2020civil]! One US physicist was so frustrated in being constantly rejected, that he added a fictitious Italian collaborator, Prof. Stronzo Bestiale from Palermo, Sicily [@penders2020civil]. Not only does Prof. Bestiale not exist, but _stronzo bestiale_ means 'massive turd' in Italian. Another German physicist Prof. Alois Kabelschacht has co-authored a number of papers, but does not exist and is actually the German for 'cable duct' - a label next to a door in the Max Planck Institute for Physics in Munich, and regularly used as a straw man in the institute [@penders2020civil]. No doubt fictitious authors abound, but these are likely to come to an end with an increase in regulation.

These days we have [ORCID](#orcid) that attempts to register all authors, turning them into numbers ([see Part III](#orcid)). One of the benefits with the ORCID system is that you can have your name the way you want to have it, even if it is not a standard western style surname [@goyesvallejos2021what]. The need for all this regulation is because publications have turned into a kind of currency for academics. And when there is currency involved, abuse quickly results in human systems followed by the need for regulation. Hence, one of the results of going transparent is that we all become registered numbers (and our dogs won't get authorship any more). Perhaps this is just the loss of an age of innocence, at least on the part of our pets!

@@

## Ghost authorship {#ghost}
Ghost authorship is when people, not included in the author line, have contributed substantially to the study [@matheson2016ghostwriting]. Note that ghostwriting outside science is usually when someone who has the talent to write, writes the ideas of someone who doesn’t for a fee, and in return the latter takes the author line. These are called "paper mills" and they are discussed in [Part I](#papermills).

@@

### Paper mills {#papermills}
The concept of a paper mill is rather different from [ghost authors](#ghost) or even salami-slicing ([Part IV](#salami)). Paper mills involve third parties, often not included on the author line, producing material for publication from scratch: i.e. companies that specialise in producing content that will pass peer review for those who want to buy authorship. The world of paper mills is particularly shady and it is not clear what sums of money are exchanged for these types of goods. A number of years ago there was evidence that first authorship on publications was for sale [@hvistendahl2013china]. What we do know is that there are benefits to those who are put on the author line, while those that produce the manuscripts generally rewrite text and pull protocols from manuscripts that are already published. Results are often images that have already been published and/or are manipulated to suit the content (see [Part IV](#fabrication)). Hence, paper mills are a systematic and deliberate manipulation of the publication process [see @teixeiradasilva2021abuse].

## Honorary authorship {#honorary}
Honorary authorship happens when people who have not contributed meaningfully to a study are included in the author line. If publications are the currency of science then you can see how being added to other people's publications increases your apparent productivity. While this might sound surprising to you, you should know that it does happen and might be more common than you think. Wisler et al. [-@wislar2011honorary] found that honorary authorship was as high as 17.6% in medical publications. It’s worth noting that this may vary between disciplines as there are various traditions in some disciplines whereby the head of a large team may always be included as an author of a paper that emerges from the team, whether or not they were involved: the White Bull effect [@kwok2005white]. In biological sciences, teams tend to be quite small with a single or rarely multiple Principal Investigators (PIs). This means that your PI will likely be directly involved with your research and therefore also an author. Imagine a very large team with multiple PIs working under a head who insists that they are an author on every publication. This could add up to hundreds of publications in a year (e.g. [Yuri T. Struchkov is currently credited with >1600 publications on Scopus](https://www-scopus-com.ez.sun.ac.za/authid/detail.uri?authorId=35427689100)), and such prolific authorship has been questioned [e.g. @rennie1994authorship]. However, there appear to be very different levels of what could be considered credible and what incredible [e.g. 25 papers a year: @wager2015too], and I respond to this, and the general question of how prolific authors are becoming [in Part IV](#publishmore).

In the biological sciences, the area of molecular phylogenetics has traditionally honoured those who collect tissue samples with authorship on the resulting phylogenies when published. As I mentioned above, this is not always equal and has also been used to politically [honour](#honorary) or [ghost](#ghost). The reason given for honouring contributors in this way is that the studies could not have been done without the tissues. On the other side, some people have long lists of publications based on tissue donations and very little else. What is needed is transparency.

## A need for transparency - DORA {#DORA}
Given the problems with both ghost and honorary authorship, there is clearly a need for transparency about who the authors are and what they have actually contributed. This was recognised by the [Declaration of Research Assessment](DORA: https://sfdora.org/) (Figure \@ref(fig:DORA)), which has a growing number of signatories as well as some solid ideas on the way forward in assessment of research and researchers. In particular, DORA is against the use of Impact Factors ([see below](#impactfactors)) and other journal based metrics, and instead assesses the research on its own merits. DORA also encourages everyone to embrace the opportunities offered by online publication, including colour figures and unencumbered word lengths. DORA also encourages specific information to be published about individual author contributions. In short, DORA stands for transparency and it would be worth you looking at their statement and finding out whether your institution is a signatory.

(ref:DORA) **The Declaration of Research Assessment: DORA.** Push back against journal metrics, such as Impact Factor, by signing the DORA.

```{r DORA, echo=FALSE, out.width = '50%',  fig.cap="(ref:DORA)", fig.align='left', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html'), fig.link='https://sfdora.org/read/'}
knitr::include_graphics('figures/Dorabadge5.png', dpi = NA)
```

@@

## Who should be an author?
There are an increasing number of journals that now give clear instructions on who qualifies as author on a paper, and these have been formalised by the International Committee of Medical Journal Editors (ICMJE), and this area has been rapidly evolving [@baskin2011honorary]. Although these approaches are clearly made to prevent [honorary authorships](#honorary), it is hard to see how they help include [ghosted authors](#ghost), or exclude the White Bull [@kwok2005white]. Scientists are notorious in their abilities to make _post hoc_ rationalisations.

As an early career researcher, I'd suggest that you use questions around authorship as a way to formulate how participating researchers fit into the study. By openly attributing roles to a checklist, everyone buys into what is expected of them. Any additions to this author line-up should be openly discussed amongst all authors, and attributed to the checklist, as the project proceeds.

### 2021-12-15 1.07-citations.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

Citations matter as they reflect the number of times people who found your research useful. Although you can look at the citations of each paper, your publications will most likely be summarised into a citation metric to get an overview of how your work is used. Here I provide a guide to some of these [metrics](#metrics), how they are calculated and used. Growing your citations by getting your research cited is not always straightforward. Publishing your research in journals with higher [Impact Factors](#impactfactor) will likely help you get your research read, but there are lots of other tactics and I give some suggestions for [increasing your citations](#increasingcitations).

## What are your citation metrics? {#metrics}
A number of different databases compile metrics which you can access. In addition, they compile a bunch of extra performance metrics that you would have to pay to access, and in general these are available to a select number of recruitment staff at your institution. Here I will talk about Google Scholar, Scopus and the Web of Science [see @martin-martin2018google; @martin-martin2021google on their differences]. In summary, for Biological Sciences, the Web of Science (WoS) and scopus are likely to give you very similar statistics. Scopus favours social-sciences a little more, so if your subject area does cross-over then you might see some extra citations there. Google Scholar is going to give you the highest metrics, but it doesn't mean that they are particularly false (although this is certainly the not the best curated database). The joy of Google Scholar is that it is much more cosmopolitan on the journals and theses that it includes. So it actually gives you a better idea of who is using your research. It will also pick up grey literature and predatory publishers, so you should interpret the outputs with care.

Each of these databases offers you the opportunity of curating your own author profile that (in the case of Scopus and WoS) will link with your ORCID, and (in the case of WoS) link with your Publons account. Your institution may want you to actively curate your profile to help them with assessing your institution's metrics. It is surprising how bad these are when left to the bots, especially if you have your name recorded in more than one way. I also think that it's worthwhile creating and curating a Google Scholar profile, and link it directly to your [website](#website). As an editor, this is one of the easiest ways of looking up an author or potential reviewer. For anyone that uses Google Scholar, your name will then be underlined, and they can quickly find your other published works. Google Scholar also indexes preprints, which may be a boon to you as an Early Career Researcher.

@@

Your H-index will grow very quickly as you publish articles from your thesis, but the growth slows as the H-index grows as it requires more citations of particular papers (of your top cited papers). Growing an H-index is therefore very difficult, and it's probably one of the hardest of these metrics to manipulate. Because the H-index is cumulative, people with very large H-indices (>100) are usually very senior academics with very good [networks](#networks), and low R numbers.

Note that while the H-index is usually calculated for individuals, but groups of people can also have an H-index, such as a department, a school, faculty or even university. Indeed, you can calculate an H-index for any set of articles using the citations that they receive.

#### Normalisaing the H-index and other citation metrics
The H-index (and other indices) are sometimes normalised for field of study (fixed - dynamic; broad - narrow), types of citing documents and year of publication [see @ioannidis2016citation]. You can expect to see some of these field normalisations, for example by [Dimensions](https://badge.dimensions.ai/) (see Figure \@ref(fig:Dimensions)) who provide you with, total citations (as the name implies), Recent citations (citations within the last 2 years), field citation ratio (comparing the article to ones of a similar age in in a similar field), and the Relative citation ratio (relative to 1.0, this gives you an idea of how well cited the article is compared to others in its area of research). As an Early Career Researcher, you may feel that your H-index is not particularly impressive (Table \@ref(tab:cite)), in which case, you may want to consider some of these normalised metrics if they show your articles in a more favourable light.

@@

@@

@@

At this point it is worth telling the tale of Neil Hall, who in 2014 wrote a slightly mischievous short article proposing a new citation metric which he called the **K-index** to reflect the number of followers on Twitter divided by the total number of citations of the same researcher [@hall2014kardashian]. In the article, he found (with a non-random selection of Twitter researchers) that there was indeed a correlation between well-cited individuals and the number of their Twitter followers. He then postulated that some people with large numbers of Twitter followers (high celebrity) were not authoritative as they did not have lots of citations (a measure of scientific value):

@@

You should not add citations to manuscripts that have simply been submitted to other journals. If you really need to, then submit and cite as a [preprint](#preprint) is really your only option. Some editors will not accept citations of preprints. An exception to this might be when papers are submitted together to the same journal at the same time. If this is the case, then you can point this out to the editor in your [cover letter](#letter).

@@

### Presenting the results of your paper at a confrence {#conference}
Probably the most traditional way of generating a bigger audience for your work is to talk about it (or present it as a poster) at a conference with an audience that is likely to appreciate your work. Everyone else at the conference will likely also be wanting you to know their work, so you'll need to target your audience carefully (possibly in a [symposium](#symposium)), and make sure that the audience can access your work (for example through a QR code in your slide or on the poster). Making your talk memorable will likely have people remember your work, and cite it if they are undertaking a similar or relevant study.

@@

### Not so legitimate ways of increasing citations {#notlegitimate}
As well as the legitimate ways of increasing citations and visibility of your work that I've listed above there are other illegitimate ways that I have heard about. I have heard of certain laboratories that have _quid pro quo_ arrangements where they agree to cite each other's publications [see @ritchie2020science].

Another way of increasing citations might be to have a co-author that is very well known. Certainly having a co-author that is very well known is likely to increase your citations but only if they actually do something to contribute to the work. See the section on [honorary authorship](#honorary).

@@

Journals have also been known to manipulate citations in order to increase their [Impact Factors](#impactfactor) and therefore their perceived level of quality. A group of physics journals from Romania were found to have clearly manipulated self-citations to increase their own impact factors [see @heneberg2016excessive]. There can also be a level of coercive pressure from journal editors to cite papers from their journals published within the two year citation window [@chorus2016large-scale]. Such practices are now systematically analysed by the larger literature databases, and clear levels of [Impact Factor](#impactfactor) manipulation result in delisting of the journal from the database.

@@

>>>>

Citations matter as they reflect the number of times your research was found to be useful. Although you can look at the citations of each paper, your publications will most likely be summarised into a citation metric to get an overview of how your work is used. Here I provide a guide to some of these [metrics](#metrics), how they are calculated and used. Throughout this book, I suggest that metrics should not be used to measure the [performance of academics](#DORA), and later how the [gamification](#badscience) of metrics has led to the detriment of the [scientific project](#closedtoopen) [see @chapman2019games]. However, this does not mean that you should not be interested in the dissemination of your work, having it read as widely as possible, and ultimately cited when appropriate.

Publishing your research in journals with higher [Impact Factors](#impactfactor) will likely help you get your research read by a larger audience. But you can increase your own readership, and this chapter goes through some of ways in which you can disseminate your work and thereby [increase your citations](#increasingcitations) of your work.

## What are your citation metrics? {#metrics}
A number of different databases compile metrics which you can access. In addition, they compile a bunch of extra performance metrics that you would have to pay to access, and in general these are available to a select number of recruitment staff at your institution. Here I will talk about Google Scholar, Scopus and the Web of Science [see @martin-martin2018google; @martin-martin2021google on their differences]. In summary, for Biological Sciences, the Web of Science (WoS) and Scopus are likely to give you very similar statistics. Scopus favours social-sciences a little more, so if your subject area does cross-over then you might see some extra citations there. Google Scholar is going to give you the highest metrics, but it doesn't mean that they are particularly false (although this is certainly the not the best curated database). The joy of Google Scholar is that it is much more cosmopolitan on the journals and theses that it includes. So it actually gives you a better idea of who is using your research. It will also pick up grey literature and predatory publishers, so you should interpret the outputs with care.

Each of these databases offers you the opportunity of curating your own author profile that (in the case of Scopus and WoS) will link with your ORCID, and (in the case of WoS) link with your Publons account. Your institution may want you to actively curate your profile to help them with assessing your institution's metrics. It is surprising how bad these are when left to the bots, especially if you have your name recorded in more than one way. Use open source tools where you can, and beware of linking too many publisher owned tools together as these can also be used to [spy on you](#capture). I also think that it's worthwhile creating and curating a Google Scholar profile, and link it directly to your [website](#website). As an editor, this is one of the easiest ways of looking up an author or potential reviewer. For anyone that uses Google Scholar, your name will then be underlined, and other users can quickly find your other published works. Google Scholar also indexes preprints, which may be a boon to you as an Early Career Researcher.

@@

Your H-index will grow very quickly as you publish articles from your thesis, but the growth slows as the H-index grows as it requires more citations of particular papers (of your top cited papers). Growing an H-index is therefore very difficult, and it's probably one of the hardest of these metrics to manipulate. Because the H-index is cumulative, people with very large H-indices (>100) are usually very senior academics with very good [networks](#networks), and low [R numbers](#r-index).

Note that while the H-index is usually calculated for individuals, groups of people can also have an H-index, such as a department, a school, faculty or even university. Indeed, you can calculate an H-index for any set of articles using the citations that they receive.

#### Normalisaing the H-index and other citation metrics {#normal}
The H-index (and [other indices](#5year)) are sometimes normalised for field of study (fixed - dynamic; broad - narrow), types of citing documents and year of publication [see @ioannidis2016citation]. You can expect to see some of these field normalisations, for example by [Dimensions](https://badge.dimensions.ai/) (see Figure \@ref(fig:Dimensions)) who provide you with, total citations (as the name implies), Recent citations (citations within the last 2 years), field citation ratio (comparing the article to ones of a similar age in in a similar field), and the Relative citation ratio (relative to 1.0, this gives you an idea of how well cited the article is compared to others in its area of research). As an Early Career Researcher, you may feel that your H-index is not particularly impressive (Table \@ref(tab:cite)), in which case, you may want to consider some of these normalised metrics if they show your articles in a more favourable light.

@@

### m-index
Another idea for decreasing the impact of older researchers is to consider the m-index [@hirsch2005index]. This takes the standard H-index, and divides it by the number of years since the first publication of that researcher. The m-index has a levelling effect between older and younger researchers, with high scores being ~1.5 [@kelly2006h]. However, although the m-index might be a better way of comparing young and old researchers, as with other indices, reducing careers to single numbers should always be considered dubious [@kelly2006h]. Moreover, such methods clearly place people who have a hiatus in their career, for example to raise a family, at a disadvantage.

### H-frac
H-frac is the same as the H-index, except that citations for articles are divided by the number of authors [@koltun2021h-index]. Koltun and Hafner [-@koltun2021h-index] propose that H-frac is more robust as it represents a better correlation with awards to individual scientists than the H-index (see their [excellent website](https://h-frac.org/) for an explanation). Their analysis shows that it accounts for changes in author behaviour, such as [hyperauthorship](#authors), based on their retrospective analysis of authors in physics. I am wary of advocating a new index that may drive future behaviour in ascribing [authorship](#author). If citations are divided by number of authors, this will likely make lead authors more circumspect about adding potential authors to their byline. Given that there are already difficulties in some [authors](#authors) being represented, I prefer not to have a disincentive for adding authors. Like other analyses that are made on vast amounts of existing data in [citation databases](http://www.howtowriteaphd.org/databases.html), this does not predict how future behaviour will change when the incentive changes.

@@

### Open Access
The subject of Open Access (OA) is covered in [Part II](#openaccess1), and so won't be covered in detail here. It is important only to mention here that having your article OA will likely increase citations [e.g. @piwowar2018state], but that you don't need to pay a publisher in order to produce an OA version of your article.

### Open Data and other open resources in repositories
There are studies that show that [transparent practices](#transparency2) such as providing your data [e.g. @piwowar2007sharing] and code on an Open Access platform (like  [Bitbucket](https://bitbucket.org/product), [Github](https://github.com/), [OSF](https://osf.io/) and [Zenodo](https://zenodo.org/)) increases citations not only to your manuscript, but also to the DOI associated with each of these repositories [e.g. @fernandez-juricic2021why]. Making transparency a part of all your research will also make your work more reproducible, a trait that we hope will be considered over and above other metrics mentioned here.

@@

At this point it is worth telling the tale of Neil Hall, who in 2014 wrote a slightly mischievous short article proposing a new citation metric which he called the **K-index** to reflect the number of followers on Twitter divided by the total number of citations of the same researcher [@hall2014kardashian]. In the article, he found (with a non-random selection of 40 researchers on Twitter) that there was indeed a correlation between well-cited individuals and the number of their Twitter followers. He then postulated that some people with large numbers of Twitter followers (high celebrity) were not authoritative as they did not have lots of citations (a measure of scientific value):

@@

You should not add citations to manuscripts that have simply been submitted to other journals. If you really need to, then submitting and citing as a [preprint](#preprint) is really your only option. Some editors will not accept citations of preprints. An exception to this might be when papers are submitted together to the same journal at the same time. If this is the case, then you can point this out to the editor in your [cover letter](#letter).

@@

### Presenting the results of your paper at a conference {#conference}
Probably the most traditional way of generating a bigger audience for your work is to talk about it (or present it as a poster) at a conference with an audience that is likely to appreciate your work. Everyone else at the conference will likely also be wanting you to know their work, so you'll need to target your audience carefully (possibly in a [symposium](#symposium)), and make sure that the audience can access your work (for example through a QR code in your slide or on the poster). Making your talk memorable will likely have people recall your work, and cite it if they are undertaking a similar or relevant study.

@@

### Not so legitimate ways of increasing citations {#notlegitimate}
In addition to the legitimate ways of increasing citations and visibility of your work that I've listed above there are other illegitimate ways that I have heard about. I have heard of certain laboratories that have _quid pro quo_ arrangements where they agree to cite each other's publications [see @ritchie2020science].

Another way of increasing citations might be to have a co-author who is very well known. Certainly having a co-author who is very well known is likely to increase your citations but only if they actually do something to contribute to the work. See the section on [honorary authorship](#honorary).

@@

Journals have also been known to manipulate citations in order to increase their [Impact Factors](#impactfactor) and therefore their perceived level of quality. A group of physics journals from Romania were found to have clearly manipulated self-citations to increase their own impact factors [see @heneberg2016excessive]. There can also be a level of coercive pressure from journal editors to cite papers from their journals published within the two year citation window [@chorus2016large-scale]. Editors can also form citation cartels in order to improve the rankings of their journals [@franck1999scientific]. Such practices are now systematically analysed by the larger literature databases, and clear levels of [Impact Factor](#impactfactor) manipulation result in delisting of the journal from the database.

@@

Interestingly, there is some evidence that a similar effect is happening in journals, with those that have the highest [Impact Factors](#impactfactor) inflating at faster rates than those with low Impact Factor [@neff2010not]. This is of concern, as it suggests that there will be a run-away effect of journals with top Impact Factors - something only of interest to the publishing houses that own them, and prompting [bad science](#badscience).
